---
title: "The Effect of Vitamin C on Tooth Growth in Guinea Pigs"
author: "Aseem Rohatgi"
output: pdf_document
---

```{r setup, include=FALSE}
library(dplyr)
library(ggplot2)
```

# The Data

The dataset we'll be using comes pre-built in R. It's stored as a data frame under the name `ToothGrowth`. The documentation for this dataset tells us that it is a collection of three variables-one response variable stating the length of odontoblasts in 60 guinea pigs, and two explanatory variables-one representing the method of delivery of vitamin C, and the other representing the dosage. First we look at its structure.

```{r structure}
str(ToothGrowth)
```

We see that the variable `dose`, which represent the three levels of dosage is stored as a numeric variable. We should keep this in mind if we are to use it as a factor later on in our analysis.
We should also check how the sample sizes of these groups differ.

```{r sample size}
ToothGrowth %>% count(supp, dose)
```

The sample sizes are all the same, albeit small. Let's now look at some basic summary statistics for this data set, with and without grouping.

```{r summary EDA, echo=F}
ToothGrowth %>% group_by(dose, supp) %>%
  summarise(quartile1=quantile(len,0.25), median=median(len),
            quartile2=quantile(len,0.75), iqr= IQR(len), mean=mean(len), 
            .groups="drop")

ToothGrowth %>% 
  summarise(quartile1=quantile(len,0.25), median=median(len),
            quartile2=quantile(len,0.75), iqr= IQR(len), mean=mean(len))
```

Plotting boxplots for the different dosages in the two delivery methods would allow us to understand this data better.

```{r boxplot EDA, echo=F}
ToothGrowth %>% ggplot(aes(factor(dose), len, fill=supp)) + geom_boxplot()
```

